Categorical Predictor Embeddings in Actuarial Modeling

### 2021-04-04 manuscript/manuscript.Rmd
some typos

The working example for our experiments is as follows: Given a set of claim characteristics, we predict the losses paid 
on the property coverage of the policy. The data we use comes from the National Flood Insurance Program (NFIP) and is 
made available by the OpenFEMA initiative 
of the Federal Emergency Management Agency (FEMA)[^9]. Two datasets are made available by OpenFEMA: A policies dataset 
with exposure information, and a claims dataset with claims transactions, including paid amounts. Because there is no 
way to associate records of the two datasets, we are limited to fitting severity models on the claims dataset. While 
the 
complete dataset contains over two million transactions, for the purposes of our experiments, we limit ourselves to 
data from 2000 to August 2019 and further subsample 100K data points to make experiments feasiable on a CPU.
 The dataset can be downloaded from 
Cellar[^10] The dataset includes a rich variety of variables, from occupancy type to coarse coordinates. For our 
models, 
we work with a few selected variables that represent continuous and discrete variables of low and high cardinalities, 
which we list in Table \@ref(tab:variables)

[^9]: https://www.fema.gov/about/reports-and-data/openfema
[^10]: https://cellar.kasa.ai

@@

For our experiments, we set up a 5-fold cross validation scheme that is applied to each of the models
we introduce. In Sections \ref{models} and \ref{interpretations}, the tables and figures are based on a single fold of 
the cross validation split. In Section \ref{training-inference}, we provide the cross-validated performance
metrics of each of the models.

@@

1. A GLM with gamma distribution and log link function,
2. A neural network with unidimensional categorical embeddings,
3. A GLM with the categorical predictors replaced by the trained embeddings from Model 2,
4. A neural network with multidimensional categorical embeddings.

@@

While this paper focuses on embeddings in neural networks, we begin with a GLM to provide a common frame of 
reference since most actuaries are familiar with the technique. 
Although GLMs are commonplace and well-studied in the actuarial literature, there are still plenty of decisions to 
be made in the modeling process; see, for example, @Goldburd for an in-depth discussion. For our purpose of 
establishing a baseline, we proceed with what we perceive as reasonable decisions around feature engineering and model structure, outlined below:

@@

In Table \@ref(tab:relativities), we exhibit an exerpt of the relativities, or exponentiated fitted coefficients, 
of a couple categorical variables. Recall that the choice of the log link gives a multiplicative structure, so the
interpretation of 1.37 for `occupancy_type = "Non-residential building"` is that, all else equal, the expected 
loss for
a policy for a non-residential building is 1.37 times that of a policy for a single family home, which is the base,
or reference, level for the factor.

@@

For Model 2, we build a feedforward neural network, whose architecture is shown in Figure \@ref(fig:nn-1). 
The categorical inputs go through one-dimensional embedding
layers where they are each mapped to a scalar. The embeddings are then concatenated with the numeric predictors, which
have been normalized in data pre-processing, before being passed through a feedforward layer (with 8 hidden units 
and ReLU activation)
 to obtain a scalar
value between 0 and 1, as constrained by a sigmoid output activation. This scalar output represents the proportion of
coverage amount that is paid out, so we multiply it by the unnormalized dollar amount of the coverage amount to obtain
our prediction of the claims paid, which we evaluate using the mean squared error (MSE) loss function.

@@

With a single hidden layer, this architecture is relatively simple; without the hidden layer and
corresponding ReLU activation, we would actually recover a logistic regression structure. 
The choice of one for the embedding dimension is also due to simplicity: it is easily interpretable as the 
representation
of each factor level as a point on the real number line, also, the trained embeddings are easily incorporated into a GLM
which we will see in Model 3. In practice, the embedding dimension is a hyperparameter one can tune, for example via
cross validation, and can vary for each variable. However, we also note that the choice of one for the 
embedding output dimension is not an unreasonable one, especially when the cardinalities of the categorical variables
are not unusually large.

In Tables \@ref(tab:embeddings-occupancy-type) and \@ref(tab:embeddings-flood-zone), we show the learned embeddings
for the `occupancy_type` and an excerpt of them for `flood_zone`. While it may be tempting to compare these values
to those in Table \@ref(tab:relativities), they represent fundamentally different concepts. It is helpful to think
of the embedding values for each variable as a continuous represention of the variable itself, rather than a 
"coefficient" associated with the variable. Note that it is not straightforward to inspect the embeddings from a
table and infer relationships to the response variable, even in a shallow network such as ours, due to nonlinearities
within the model. Also note that we now have a way to incorporate the full granularity of the `flood_zone` variable,
which is why more levels are shown in Table \@ref(tab:embeddings-flood-zone) as compared to Table 
\@ref(tab:relativities). In Section \ref{interpretation}, we will discuss techniques to visualize these embeddings.

>>>>

The working example for our experiments is as follows: Given a set of claim characteristics, we predict the losses paid 
on the property coverage of the policy. The data we use comes from the National Flood Insurance Program (NFIP) and is 
made available by the OpenFEMA initiative 
of the Federal Emergency Management Agency (FEMA)[^9]. Two datasets are made available by OpenFEMA: A policies dataset 
with exposure information, and a claims dataset with claims transactions, including paid amounts. Because there is no 
way to associate records of the two datasets, we are limited to fitting severity models on the claims dataset. While 
the 
complete dataset contains over two million transactions, for the purposes of our experiments, we limit ourselves to 
data from 2000 to August 2019 and further subsample 100K data points to make experiments feasible on a CPU.
The dataset includes a rich variety of variables, from occupancy type to flood zone. For our 
models, 
we work with a few selected variables that represent continuous and discrete variables of low and high cardinalities, 
which we list in Table \@ref(tab:variables)

[^9]: https://www.fema.gov/about/reports-and-data/openfema

@@

For our experiments, we set up a 5-fold cross validation scheme that is applied to each of the models
we introduce. In Sections \ref{models} and \ref{interpretation}, the tables and figures are based on a single fold of 
the cross validation split. In Section \ref{training-inference}, we provide the cross-validated performance
metrics of each of the models.

@@

1. A GLM with gamma distribution and log link function,
2. A neural network with one-dimensional categorical embeddings,
3. A GLM with the categorical predictors replaced by the learned embeddings from Model 2,
4. A neural network with multidimensional categorical embeddings.

@@

While this paper focuses on embeddings in neural networks, we begin with a GLM to provide a common frame of 
reference, since most actuaries are familiar with the technique. 
Although GLMs are commonplace and well-studied in the actuarial literature, there are still plenty of decisions to 
be made in the modeling process; see, for example, @Goldburd for an in-depth discussion. For our purpose of 
establishing a baseline, we proceed with what we perceive as reasonable decisions around feature engineering and model structure, outlined below:

@@

In Table \@ref(tab:relativities), we exhibit an exerpt of the relativities, or exponentiated fitted coefficients, 
of a couple categorical variables. Recall that the choice of the log link gives a multiplicative structure, so the
interpretation of 1.34 for `occupancy_type = "Non-residential building"` is that, all else equal, the expected 
loss for
a policy for a non-residential building is 1.34 times that of a policy for a single family home, which is the base,
or reference, level for the factor.

@@

For Model 2, we build a feedforward neural network, whose architecture is shown in Figure \@ref(fig:nn-1). 
The categorical inputs go through one-dimensional embedding
layers where they are each mapped to a scalar. The embeddings are then concatenated with the numeric predictors, which
have been normalized in data pre-processing, before being passed through a feedforward layer (with 8 hidden units 
and ReLU activation)
 to obtain a scalar
value between 0 and 1, as constrained by a sigmoid output activation. This scalar output represents the proportion of
coverage amount that is paid out, which we then multiply by the dollar amount of the coverage amount to obtain
our prediction of the claims paid. For optimization, we use the mean squared error (MSE) loss function.

@@

With a single hidden layer, this architecture is relatively simple; without the hidden layer and
corresponding ReLU activation, we would actually recover a logistic regression structure. 
The choice of one for the embedding dimension is also due to simplicity: it is easily interpretable as the 
representation
of each factor level as a point on the real number line; also, the trained embeddings are easily incorporated into a GLM
which we will see in Model 3. In practice, the embedding dimension is a hyperparameter one can tune, for example via
cross validation, and can vary for each variable. However, we also note that the choice of one for the 
embedding output dimension is not an unreasonable one, especially when the cardinalities of the categorical variables
are not large (i.e., when they are in the 10s).

In Tables \@ref(tab:embeddings-occupancy-type) and \@ref(tab:embeddings-flood-zone), we show the learned embeddings
for `occupancy_type` and an excerpt of them for `flood_zone`. While it may be tempting to compare these values
to those in Table \@ref(tab:relativities), they represent fundamentally different concepts. It is helpful to think
of the embedding values for each variable as a continuous represention of the variable itself, rather than a 
"coefficient" associated with the variable. Note that it is not straightforward to inspect the embeddings from a
table and infer relationships to the response variable, even in a shallow network such as ours, due to nonlinearities
within the model. Also note that we now have a way to incorporate the full granularity of the `flood_zone` variable,
which is why more levels are shown in Table \@ref(tab:embeddings-flood-zone) as compared to Table 
\@ref(tab:relativities). In Section \ref{interpretation}, we will discuss techniques to visualize these embeddings.
